---
title: "Homework 4 - Programming in Base R"
format: pdf
editor: visual
---

## Task 1: Conceptual Questions

### Question 1: What is the purpose of the lapply() function? What is the equivalent purrr function?

The purpose of the lapply() function is to apply a function across different elements of a list. The regular apply() function only works with arrays, but lapply() allows us to work with lists that contain heterogeneous objects. The equivalent purrr function is

### Question 2: Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns are numeric). We want use lapply() to run the code cor(numeric_matrix, method = "kendall") on each element of the list. Write code to do this below! (I'm really trying to ask you how you specify method = "kendall" when calling lapply())

The code to do this is is lapply(X = my_list, FUN = cor, method = "kendall")

### Question 3: What are two advantages of using purrr functions instead of the BaseR apply family?

Two advantages of using purrr functions instead of the BaseR apply family is that the purrr functions are more consistent across the board, whereas there are several differences in the BaseR apply family in terms of capitalization of terms (fun and FUN) and other minor differences. The purrr functions are also cleaner and easier to read than the apply family, and while they can be tough to master at first can allow for quicker and more readable code.

### Question 4: What is a side-effect function?

A side effect function is a function that doesn't transform the data that we're working with, but instead creates something new from that data. An example of a side-effect function is plot() as we are taking the data and plotting it, but not changing anything with the data itself.

### Question 5: Why can you name a variable sd in a function and not cause any issues with the sd function?

You can name a variable sd in a function and not cause any issues with the sd function because when we create functions we essentially also create a separate environment. Functions keep created objects separate, so naming somethinf sd within a function doesn't affect sd outside the function, as the function object that was created is not kept after the function is run.

## Task 2: Writing R Functions

### Question 1: RMSE

When we start doing machine learning later in the course, a common metric used to evaluate predictions is called Root Mean Square Error (RMSE). Write a basic function (call it getRMSE()) that takes in a vector of responses and a vector of predictions and outputs the RMSE. If a value is missing for the vector of responses (i.e. an NA is present), allow for additional arguments to the mean() function (ellipses) that removes the NA values in the computation.

```{r}
getRMSE <- function(responses, predictions, ...){
  RMSE <- sqrt(mean(((responses - predictions)^2), ...))
  RMSE
}
```

### Question 2: Testing RMSE function

Test your RMSE function using this data below. Repeat after replacing two of the response values with missing values (NA_real_). Test your RMSE function with and without specifying the behavior to deal with missing values.

```{r}
#Testing data
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

#Initial test
getRMSE(resp, pred)

#Replacing 2 response values with missing values
resp[1:2] <- NA_real_

#New test, with and without specification for NA
getRMSE(resp, pred)
getRMSE(resp, pred, na.rm = TRUE)
```

### Question 3: MAE

Another common metric for evaluating predictions is mean absolute deviation given by the MAE. Write a function called getMAE() that follows the specifications of the getRMSE() function.

```{r}
getMAE <- function(responses, predictions, ...){
  MAE <- mean(abs(responses - predictions), ...)
  MAE
}
```

### Question 4: Testing MAE function

Test your MAE function using this data below. Repeat after replacing two of the response values with missing values (NA_real\_). Test your MAE function with and without specifying the behavior to deal with missing values.

```{r}
#Testing data
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

#Initial test
getMAE(resp, pred)

#Replacing 2 response values with missing values
resp[1:2] <- NA_real_

#New test, with and without specification for NA
getMAE(resp, pred)
getMAE(resp, pred, na.rm = TRUE)
```

### Question 5: Creating wrapper function

Let's create a wrapper function that can be used to get either or both metrics returned with a single function call. Do not rewrite your above two functions, call them inside the wrapper function (we would call the getRMSE() and getMAE() functions helper functions). When returning your values, give them appropriate names.

-   The function should check that two numeric (atomic) vectors have been passed (consider is.vector(), is.atomic(), and is.numeric()). If not, a message should print and the function should exit.

-   The function should return both metrics by default and include names. The behavior should be able to be changed using a character string of metrics to find.
```{r}
vec_wrapper <- function(responses, predictions, metrics = c("RMSE", "MAE"), ...){
  if(!is.vector(responses) | !is.numeric(responses) | !is.atomic(responses)){
    stop("Error! Response vector not a numeric (atomic) vector!")
  }
  else if(!is.vector(predictions) | !is.numeric(predictions) | !is.atomic(predictions)){
    stop("Error! Predictions vector not a numeric (atomic) vector!")
  }
}
```


### Question 6: Testing wrapper function

Test your new function using this data. Call it once asking for each metric individually and once specifying both metrics. Repeat with replacing two of the response values with missing values (NA_real_). Finally, test your function by passing it incorrect data (i.e. a data frame or something else instead of vectors)
```{r}
#Testing data
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

vec_wrapper(resp, pred)
```

